---
title: "Regression Exercise #1"
author: "Matthew Lee"
date: "7/27/2020"
output: 
  html_document:
    includes:
      in_header: header.HTML
---
<script type="text/javascript" src="https://cdnjs.cloudflare.com/ajax/libs/iframe-resizer/3.5.16/iframeResizer.min.js"></script>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
<center>
{{% tweet 1156589945676554241 %}} 
</center>

The following exercises will help illustrate concepts on how two variables are related, as demonstrated by their covariance and correlation. Feel free to run your own `R` code on your local machine, or utilize the code blocks on this site. You can use 
<kbd class="keyboard-key nowrap" style="border: 1px solid #aaa; -moz-border-radius: 0.2em; -webkit-border-radius: 0.2em; border-radius: 0.2em; -moz-box-shadow: 0.1em 0.1em 0.2em rgba(0,0,0,0.1); -webkit-box-shadow: 0.1em 0.1em 0.2em rgba(0,0,0,0.1); box-shadow: 0.1em 0.1em 0.2em rgba(0,0,0,0.1); background-color: #f9f9f9; background-image: -moz-linear-gradient(top, #eee, #f9f9f9, #eee); background-image: -o-linear-gradient(top, #eee, #f9f9f9, #eee); background-image: -webkit-linear-gradient(top, #eee, #f9f9f9, #eee); background-image: linear-gradient(to bottom, #eee, #f9f9f9, #eee); color: #000; padding: 0.1em 0.3em; font-family: inherit; font-size: 0.85em;">Ctrl</kbd> /
<b> <span class="nounderlines" style="border: 1px solid #ddd; background-color: #fdfdfd; padding: 1px 1px;">⌘</span></b> `+`
<kbd class="keyboard-key nowrap" style="border: 1px solid #aaa; -moz-border-radius: 0.2em; -webkit-border-radius: 0.2em; border-radius: 0.2em; -moz-box-shadow: 0.1em 0.1em 0.2em rgba(0,0,0,0.1); -webkit-box-shadow: 0.1em 0.1em 0.2em rgba(0,0,0,0.1); box-shadow: 0.1em 0.1em 0.2em rgba(0,0,0,0.1); background-color: #f9f9f9; background-image: -moz-linear-gradient(top, #eee, #f9f9f9, #eee); background-image: -o-linear-gradient(top, #eee, #f9f9f9, #eee); background-image: -webkit-linear-gradient(top, #eee, #f9f9f9, #eee); background-image: linear-gradient(to bottom, #eee, #f9f9f9, #eee); color: #000; padding: 0.1em 0.3em; font-family: inherit; font-size: 0.85em;">↵ Enter</kbd> 
to evaluate all the code in a given chunk. 



## Creating random variables

------------------------------------------------

`R` let's you generate your own random variables from common distribution functions like the normal, binomial, poisson, and uniform distributions we discussed in class.

For example, the following code generates 3 realizations/random samples from a normal distribution with mean 45 (`mean`) and standard deviation (`sd`) 19. The function `rnorm()` takes in these arguments, along with the number of draws you want (`n`) and returns `n` random draws from the normal distribution you specify. 

<!--- LEARNR EXAMPLE 1 --->
<iframe style="margin:0 auto; min-width: 100%;" id="normalEx" class="interactive" src="https://leem26.shinyapps.io/regressionEx1-normalEx/" scrolling="no" frameborder="no"></iframe>
<!--- END LEARNR EXAMPLE 1 --->


Next, generate a vector of 100 realizations from your favorite distribution and save the results to a vector (call it whatever you want!). Then, plot a histograms of all 100 realizations. This should resemble the distribution you chose to a degree. Feel free to increase the number of realizations (say from `n = 100` to `n = 1000`) and see what happens to the histogram.

<!--- LEARNR EXAMPLE 2 --->
<iframe style="margin:0 auto; min-width: 100%;" id="normalPlot" class="interactive" src="https://leem26.shinyapps.io/regressionEx1-normalPlot/" scrolling="no" frameborder="no"></iframe>
<!--- END LEARNR EXAMPLE 2 --->


## Covariance, correlation, and independence 

------------------------------------------------

In `R` we can calculate the correlation and covariance using the `cor()`and `cov()` functions. For example, the code below draws 1000 realizations from a normal distribution with mean 10 and standard deviation 2, and another 1000 realizations from a normal distribution with mean 8 and standard deviation 3. You can think of these as data on 1000 individuals on two variables: `x` and `y`. What do you get when you run this code? Why might that be?

<!--- LEARNR EXAMPLE 3 --->
<iframe style="margin:0 auto; min-width: 100%;" id="corcov" class="interactive" src="https://leem26.shinyapps.io/regressionEx1-corcov/" scrolling="no" frameborder="no"></iframe>
<!--- END LEARNR EXAMPLE 3 --->

What happens if we redefine `y` as a function of `x`? In other words, what happens if `y` and `x` are related by some given function $y = 10 \times x$? Obtain the correlation and covariance between these two variables with this new definition of `y`. 

<!--- LEARNR EXAMPLE 4 --->
<iframe style="margin:0 auto; min-width: 100%;" id="corcovScaled" class="interactive" src="https://leem26.shinyapps.io/regressionEx1-corcovScaled/" scrolling="no" frameborder="no"></iframe>
<!--- END LEARNR EXAMPLE 4 --->


To generate two random variables with known covariance and/or correlation we can also simulate a joint distribution, in this case the multivariate normal, using a function from the `MASS` package. Recall that a package in R is just a set of functions and objects that have been compiled into a single collection. One function from `MASS` is `mvrnorm()`, which as it suggests, returns values from the **multi-**variate normal distribution. 

Use the `mvrnorm()` function to generate 1000 samples/realizations from the multivariate normal distribution. Variable $X$ should have mean 10 and variable $Y$ should have mean 15. The variance of $X$ is 30, and the variance of $Y$ is 60. Let the covariance between $X$ and $Y$ be 40. You can read about the documentation for this function [here](https://www.rdocumentation.org/packages/rockchalk/versions/1.8.144/topics/mvrnorm){target="_blank"}, which will tell you how to specify the arguments `mu`, `Sigma`, and `n`. Also recall that the correlation $\rho$ between two variables is given by:

$$ \rho(X, Y) = \frac{\text{Cov}(X,Y)}{\sigma(X) \sigma(Y)}$$

Where $\text{Cov()}$ is the covariance, and $\sigma()$ is the standard deviation for each variable. The `sd()` function returns standard deviation values in `R`. Print the covariance and correlation matrices between $X$ and $Y$, and graph the relationship between these two variables on a 2-dimensional plane. What do you see?

<!--- LEARNR EXAMPLE 5 --->
<iframe style="margin:0 auto; min-width: 100%;" id="corcovMVN" class="interactive" src="https://leem26.shinyapps.io/regressionEx1-corcovMVN/" scrolling="no" frameborder="no"></iframe>
<!--- END LEARNR EXAMPLE 5 --->

This scatter plot tells us that $X$ and $Y$ co-occur with greater frequency around their means. If we want to visualize the **multi**-variate probability distribution, we need to extend our perspective to 3 dimensions. The `plotly` package allows us to create really neat interactive plots -- check out the one below! 


```{r, echo=TRUE, out.width ='100%', eval = TRUE, warning=FALSE, message=FALSE}
library(MASS)
library(plotly)
library(mvtnorm)
library(ggplot2)
library(dplyr)
library(widgetframe)

# generate data from the multivariate normal
mvnorm1000 <- 
  mvrnorm(
    n = 1000, 
    mu = c(10, 15), 
    Sigma = 
      matrix(c(30, 40, 40, 60), 
             nrow=2, ncol=2, byrow=FALSE)
          )

# compute density surface
dens <- kde2d(mvnorm1000[,1], mvnorm1000[,2])

# plot object
p <-
  plot_ly(x = dens$x,
          y = dens$y,
          z = dens$z) %>% 
  add_surface()

# place plot object in iframe
frameWidget(p, height = 500, width = '100%', )
```



<script>
  iFrameResize({}, ".interactive");
</script>